---
title: "longitudinal cognition data"
author: "Ruoying Deng"
date: "2024-10-08"
output: html_document
---
```{r}
#library
library(lubridate)
library(ADNIMERGE)
library(dplyr)
library(lme4)
library(ggplot2)
library(car)
library(nlme)
```



```{r}
# Convert examdate and visdate columns to Date class
cognitive_marker_date <- cognitive_marker_distinct %>%
  mutate(examdate = ymd(examdate),visdate = ymd(visdate))

# Filter out rows where the difference between examdate and visdate is more than one month
cognitive_marker_filtered <- cognitive_marker_date %>%
  filter((interval(examdate,visdate) %/% months(1)) > 1)%>%
  select(rid,viscode,phase,examdate,visdate,everything())

cognitive_marker_distinct%>%
  filter(rid %in% c(66,70,89,112,150))
#i noticed that even though there are cases where examdate before visdate, the bl,and other viscode correspond to examdate(66m12,70bl,89m24,112m06,150m24)

# Remove the filtered rows from the original dataset
#cognitive_marker_date_clean <- anti_join(cognitive_marker_date, cognitive_marker_filtered, by = c("visdate", "examdate"))

```

```{r}
biomarker<-ADNIMERGE::adnimerge
cl_biomarker <- biomark%>%
  janitor::clean_names()
```

```{r}
# we notice that some people missing baseline
cognitive_marker_distinct %>%
  group_by(rid) %>%
  summarise(has_baseline = any(viscode2 == "bl")) %>%  
  filter(has_baseline == FALSE) %>%
  pull(rid)%>%
  length()
# total 3538 patients, 1118 of them missing baseline info

# we notice some patient start screening dec 2023, therefore only one row of sc or bl exist, we filter out them for better longitudinal study
single_record_rid <- cognitive_marker_distinct %>%
  group_by(rid) %>%
  summarise(single_record = n() == 1) %>%
  filter(single_record) %>%
  pull(rid)

filtered_longitudinal_data <- cognitive_marker_distinct %>%
  filter(!rid %in% single_record_rid)

# 0 missing bl now
filtered_longitudinal_data %>%
  group_by(rid) %>%
  summarise(has_baseline = any(viscode2 == "bl")) %>%  
  filter(has_baseline == FALSE) %>%
  pull(rid)%>%
  length()

filtered_longitudinal_data%>%
  group_by(viscode2)%>%
  count()

#check on uns1
filtered_longitudinal_data%>%
  filter(rid == 168)

filtered_longitudinal_data_clean <- filtered_longitudinal_data %>%
  mutate(viscode2 = ifelse(viscode2 == "uns1", "m12", viscode2))%>%
  select(rid,viscode2,phase,examdate,phc_age_cognition,everything())

filtered_longitudinal_data_clean%>%
  filter(rid == 168)

```

```{r}
# Extract baseline examdate for each participant
baseline_dates <- filtered_longitudinal_data_clean %>%
  filter(viscode2 == "bl") %>%
  select(rid, examdate) %>%
  rename(baseline_examdate = examdate)

# Join baseline dates back to the main dataset
filtered_longitudinal_data_clean <- filtered_longitudinal_data_clean %>%
  left_join(baseline_dates, by = "rid")

filtered_longitudinal_data_clean <- filtered_longitudinal_data_clean %>%
  mutate(time_since_baseline = round(as.numeric(difftime(examdate, baseline_examdate, units = "days") / 365), 2)) %>%
  select(rid, viscode2, time_since_baseline, phase, examdate, everything())



```


```{r}
# check on four variables
variables <- c("phc_mem", "phc_exf", "phc_lan", "phc_vsp")
filtered_longitudinal_data_clean%>%
  filter(is.na(phc_diagnosis))
# Filter out missing values for relevant columns
filtered_longitudinal_data_clean_no_na <- filtered_longitudinal_data_clean %>%
  filter(!is.na(time_since_baseline) & !is.na(phc_mem) & !is.na(phc_diagnosis))

#merge APOE4 from adnimerge
adnimerge <- adnimerge %>%
  janitor::clean_names()

merged_longitudinal_data_no_na <- filtered_longitudinal_data_clean_no_na %>%
  left_join(adnimerge %>% 
  select(rid, viscode, examdate, apoe4), by = c("rid", "viscode", "examdate"))  

# Convert dataset from wide to long format and summarize the means
mean_phc_data <- merged_longitudinal_data_no_na %>%
  select(phc_diagnosis, time_since_baseline, all_of(variables)) %>%
  pivot_longer(cols = all_of(variables), names_to = "variable", values_to = "value") %>%
  group_by(phc_diagnosis, time_since_baseline, variable) %>%
  summarise(mean_value = mean(value, na.rm = TRUE), .groups = 'drop')

# Plot the longitudinal model for each variable and combine using facets
ggplot(mean_phc_data, aes(x = time_since_baseline, y = mean_value, color = as.factor(phc_diagnosis))) +
  geom_point(alpha=0.3) +
    geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(
    name = "Diagnosis",
    values = c("1" = "green", "2" = "blue", "3" = "red"),
    labels = c("1" = "Normal Cognition", "2" = "Mild Cognitive Impairment", "3" = "Alzheimer's Disease")
  ) +
  labs(
    title = "Longitudinal Model of Mean Scores by Time Since Baseline",
    x = "Time Since Baseline (Years)",
    y = "Mean Score"
  ) +
  facet_wrap(~variable, scales = "free_y") +
  theme_minimal()
```



```{r}
# Prepare data: Split age into baseline and longitudinal components
merged_longitudinal_data_no_na <- merged_longitudinal_data_no_na %>%
  group_by(rid) %>%
  mutate(
    age_baseline = phc_age_cognition[viscode2 == "bl"],   # Baseline age for each individual
  ) %>%
  drop_na(phc_ethnicity)%>%
  ungroup()

# 6351 without apoe4 and 4 without ethinicity, leave apoe4 for now 
merged_longitudinal_data_no_na%>%
  group_by(apoe4)%>%
  count()

# missing 
merged_longitudinal_data_no_na%>%
  group_by(medication_category)%>%
  count()

#Convert categorical variables to factors and define their levels explicitly
merged_longitudinal_data_no_na_clean <- merged_longitudinal_data_no_na %>%
  mutate(
    phc_diagnosis = factor(phc_diagnosis, levels = c(1, 2, 3)),
    phc_race = factor(phc_race, levels = unique(phc_race), exclude = NULL), 
   #education as continuous
    phc_ethnicity = factor(phc_ethnicity, levels = unique(phc_ethnicity)),
    phc_sex = factor(phc_sex, levels = unique(phc_sex), exclude = NULL),
  )


# Fit a linear mixed-effects model with random slope and random intercept with covariate:Phc_diagnosis +race +apoe4 from adnimerge + education + ethinicity + sex  +diagnosis by time interaction
model_lme <- lme(
  phc_mem ~ time_since_baseline * phc_diagnosis + phc_race + phc_ethnicity + phc_education + phc_sex + age_baseline,
  random = ~ time_since_baseline | rid,
  data = merged_longitudinal_data_no_na_clean,
  method = "REML")


summary(model_lme)

# CI
intervals(model_lme, which = "fixed")


model_summary <- summary(model_lme)
fixed_effects <- model_summary$tTable

# p-values based on the t-statistic and degrees of freedom
p_values <- 2 * (1 - pt(abs(fixed_effects[, "t-value"]), model_summary$fixDF$X))
fixed_effects_with_p <- cbind(fixed_effects, "p-value" = p_values)

print(fixed_effects_with_p)

fixed_effects <- fixef(model_lme)

# Extract fixed effect coefficients
intercept <- fixef(model_lme)["(Intercept)"]
slope <- fixef(model_lme)["time_since_baseline"]

plot_data <- data.frame(
  time_since_baseline = seq(0, max(merged_longitudinal_data_no_na_clean$time_since_baseline)))

# Calculate the predicted values based on the intercept and slope
plot_data$predicted_phc_mem <- intercept + slope * plot_data$time_since_baseline

ggplot() + geom_point(data = mean_phc_data, aes(x = time_since_baseline, y = mean_value), size = 1.5, alpha = 0.7) + geom_line(data = plot_data, aes(x = time_since_baseline, y = predicted_phc_mem), color = "blue", size = 1.2) + labs(title = "Fixed Effects - Model Prediction",
       x = "Time Since Baseline (Years)",
       y = "Predicted PHC Memory Score") +
  theme_minimal()


```

```{r}
# check on variable medication_category
missing_medication_category <- merged_longitudinal_data_no_na_clean %>%
  filter(is.na(medication_category))

missing_medication_category%>%
  group_by(rid)%>%
  count()%>%
  arrange(desc(n))

# Anti-join to find rows in cl_cognitive_marker that did not match in medication_with_keymed_updated
unmatched_rows <- missing_medication_category %>%
  select(rid, viscode2, phase) %>%
  anti_join(medication_with_keymed_updated %>% select(rid, viscode2, phase),
            by = c("rid", "viscode2", "phase"))

# View the unmatched rows
unmatched_rows

```


```{r}
plot_data <- merged_longitudinal_data_no_na_clean %>%
  select(phc_diagnosis, phc_race, apoe4, phc_education, phc_sex, phc_ethnicity, age_baseline, time_since_baseline) %>%
  distinct() %>%
  group_by(phc_diagnosis) %>%  # Retain grouping by each diagnosis group
  mutate(
    # Set fixed values for categorical covariates
    phc_race = ifelse(is.factor(phc_race), levels(phc_race)[1], phc_race),
    apoe4 = ifelse(is.factor(apoe4), levels(apoe4)[1], apoe4),
    phc_sex = ifelse(is.factor(phc_sex), levels(phc_sex)[1], phc_sex),
    phc_ethnicity = ifelse(is.factor(phc_ethnicity), levels(phc_ethnicity)[1], phc_ethnicity),

    # Set median value for continuous variables
    age_baseline = median(merged_longitudinal_data_no_na_clean$age_baseline, na.rm = TRUE),
    phc_education = median(merged_longitudinal_data_no_na_clean$phc_education, na.rm = TRUE)  
  ) %>%
  ungroup()

model_lme_refined <- lme(
  phc_mem ~ time_since_baseline * phc_diagnosis + phc_race + phc_ethnicity +
            phc_education + phc_sex + age_baseline,
  random = ~ time_since_baseline | rid/phc_diagnosis,# random intercepts and slopes for time_since_baseline are nested within each rid (individual ID), and rid itself is nested within phc_diagnosis groups, capturing variability both within individuals and between diagnosis groups.
  data = merged_longitudinal_data_no_na_clean,
  method = "REML",control = lmeControl(opt = "optim", maxIter = 100, msMaxIter = 100)
)
summary(model_lme_refined)

anova(model_lme, model_lme_refined)


# Predict values using the lme model for each level of phc_diagnosis
plot_data <- plot_data %>%
  mutate(predicted_phc_mem = predict(model_lme_refined, newdata = ., level = 0))

# Plotting
ggplot(plot_data, aes(x = time_since_baseline, y = predicted_phc_mem, color = as.factor(phc_diagnosis), group = phc_diagnosis)) +
  geom_point(data = mean_phc_data, aes(x = time_since_baseline, y = mean_value, color = as.factor(phc_diagnosis)), size = 1.5, alpha = 0.4) +
  geom_line(size = 1.2) +
  stat_summary(fun = mean, geom = "line", aes(group = phc_diagnosis), size = 1.5, linetype = "dashed") +
  scale_color_manual(
    name = "Diagnosis",
    values = c("1" = "green", "2" = "blue", "3" = "red"),
    labels = c("1" = "Normal Cognition", "2" = "Mild Cognitive Impairment", "3" = "Alzheimer's Disease")
  ) +
  labs(
    title = "Predicted Memory Score Over Time by Diagnosis Group",
    x = "Time Since Baseline (Years)",
    y = "Predicted Memory Score",
    color = "Diagnosis"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")


```



```{r}
ggplot() +
  geom_point(data = merged_longitudinal_data_no_na_clean, aes(x = time_since_baseline, y = phc_mem, color = as.factor(phc_diagnosis)), alpha = 0.1) +
  geom_line(data = plot_data, aes(x = time_since_baseline, y = predicted_phc_mem, color = as.factor(phc_diagnosis), group = phc_diagnosis), size = 1.5) +
  scale_color_manual(
    name = "Diagnosis",
    values = c("1" = "green", "2" = "blue", "3" = "red"),
    labels = c("1" = "Normal Cognition", "2" = "Mild Cognitive Impairment", "3" = "Alzheimer's Disease")
  ) +
  facet_wrap(~ phc_diagnosis, scales = "fixed") +
  theme_minimal() +
  theme(legend.position = "bottom")


```

```{r}
# plot using fixed effects of the model (intercept and slope)
fixed_effects <- fixef(model_lme)
print(fixed_effects)

# Extract intercept and slopes
intercept <- fixed_effects["(Intercept)"] # Baseline intercept
slope_time <- fixed_effects["time_since_baseline"] # Slope for time_since_baseline
diagnosis_effects <- fixed_effects[grep("phc_diagnosis", names(fixed_effects))]

# Generate a sequence for time_since_baseline
time_seq <- seq(0, 15, length.out = 100)

# Compute trend lines for each diagnosis group
trend_data <- data.frame(
  time_since_baseline = rep(time_seq, each = 3),
  phc_diagnosis = factor(rep(1:3, times = length(time_seq))) # Assuming Diagnosis 1, 2, 3
)

# Add fixed effect predictions for each diagnosis group
trend_data <- trend_data %>%
  mutate(
    intercept = intercept + 
                ifelse(phc_diagnosis == 2, diagnosis_effects["phc_diagnosis2"], 0) +
                ifelse(phc_diagnosis == 3, diagnosis_effects["phc_diagnosis3"], 0),
    predicted_phc_mem = intercept + slope_time * time_since_baseline
  )

ggplot() +
  geom_point(data = merged_longitudinal_data_no_na_clean, 
             aes(x = time_since_baseline, y = phc_mem, color = as.factor(phc_diagnosis)), 
             alpha = 0.4) +
  geom_line(data = trend_data, 
            aes(x = time_since_baseline, y = predicted_phc_mem, color = as.factor(phc_diagnosis), group = phc_diagnosis), 
            size = 1.2) +
  scale_color_manual(
    name = "Diagnosis",
    values = c("1" = "green", "2" = "blue", "3" = "red"),
    labels = c("1" = "Normal Cognition", "2" = "Mild Cognitive Impairment", "3" = "Alzheimer's Disease")
  ) +
  labs(
    title = "Predicted Memory Score Over Time by Diagnosis Group (Fixed Effects)",
    x = "Time Since Baseline (Years)",
    y = "Memory Score"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

```




```{r}
# participants with baseline phc_diagnosis = 1
baseline_diagnosis_1 <- merged_longitudinal_data_no_na %>%
  filter(phc_diagnosis == 1 & viscode2 == "bl") %>%
  select(rid) %>% 
  distinct()

# Check if these participants convert to phc_diagnosis = 3 at any follow-up time
conversion_to_ad <- merged_longitudinal_data_no_na %>%
  filter(rid %in% baseline_diagnosis_1$rid & phc_diagnosis == 3) %>%
  distinct(rid)

num_converted <- nrow(conversion_to_ad)

# since keep apoe4 = empty, find out more patients from 1 to 3
cat("Number of participants with baseline diagnosis 1 who converted to AD (diagnosis 3):", num_converted, "\n")

converted_data <- merged_longitudinal_data_no_na %>%
  filter(rid %in% conversion_to_ad$rid) %>%
  arrange(rid, time_since_baseline)

ggplot(converted_data, aes(x = time_since_baseline, y = phc_diagnosis, color = factor(rid))) +
  geom_line(size = 1.2) +
   geom_point(aes(shape = medication_category)) +
   scale_y_continuous(breaks = c(1, 2, 3), labels = c("Normal (1)", "MCI (2)", "AD (3)")) +
   labs(title = "Diagnosis Progression Over Time for Participants Converting to AD",
        x = "Time Since Baseline (Years)",
        y = "PHC Diagnosis",
        color = "Participant ID",
        shape = "Medication Type") +
   theme_minimal() +
   theme(legend.position = "right") +
   facet_wrap(~ rid)
```

```{r}
#data at baseline
cross_sectional_data <- merged_longitudinal_data_no_na_clean %>%
  filter(time_since_baseline == 0)

# Fit the linear model with all baseline covariates
cross_sectional_model <- lm(phc_mem ~ phc_diagnosis + phc_race + phc_education + phc_sex + age_baseline+phc_ethnicity, 
                            data = cross_sectional_data)

summary(cross_sectional_model)


# Smooth phc_mem using loess for the overall data
smoothed_loess <- loess(phc_mem ~ age_baseline, data = cross_sectional_data, span = 0.5)
cross_sectional_data$smoothed_phc_mem <- predict(smoothed_loess)

# Plot actual data points and the smoothed regression line
ggplot(cross_sectional_data, aes(x = age_baseline, y = phc_mem)) +
  geom_point(alpha = 0.5) +
  geom_line(aes(y = smoothed_phc_mem), color = "blue", size = 1.2) +
  labs(title = "Cross-Sectional Model: Smoothed PHC Memory Score by Baseline Covariates",
       x = "Age at Baseline (Years)",
       y = "PHC Memory Score") +
  theme_minimal()

# Smooth the data by phc_diagnosis group
cross_sectional_data <- cross_sectional_data %>%
  group_by(phc_diagnosis) %>%
  mutate(smoothed_phc_mem = predict(loess(phc_mem ~ age_baseline, span = 0.5))) %>%
  ungroup()

# Plot smoothed regression lines by phc_diagnosis group
ggplot(cross_sectional_data, aes(x = age_baseline, y = phc_mem, color = factor(phc_diagnosis))) +
  geom_point(alpha = 0.5) +
  geom_line(aes(y = smoothed_phc_mem), size = 1.2) +
  labs(title = "Cross-Sectional Model: Smoothed PHC Memory Score by Baseline Covariates",
       x = "Age at Baseline (Years)",
       y = "PHC Memory Score",
       color = "PHC Diagnosis") +
  theme_minimal() +
  theme(legend.position = "bottom")

```
```{r}
# Longitudinal Model Predictions
longitudinal_data <- merged_longitudinal_data_no_na_clean %>%
  mutate(predicted_phc_mem = predict(model_lme, newdata = merged_longitudinal_data_no_na_clean, level = 0))

# Combine Longitudinal Predictions with Cross-Sectional Data
combined_data <- bind_rows(
  cross_sectional_data %>%
    mutate(type = "Cross-Sectional"),
  longitudinal_data %>%
    mutate(type = "Longitudinal")
)

# Plot Cross-Sectional and Longitudinal Trends
ggplot(combined_data, aes(x = age_baseline, y = phc_mem, color = factor(phc_diagnosis), linetype = type)) +
  geom_point(alpha = 0.5) +
  geom_line(aes(y = smoothed_phc_mem), size = 1.2) +
  labs(title = "Cross-Sectional vs. Longitudinal Trends in PHC Memory Scores",
       x = "Age at Baseline (Years)",
       y = "PHC Memory Score",
       color = "PHC Diagnosis",
       linetype = "Model Type") +
  theme_minimal() +
  theme(legend.position = "bottom")

```


```{r}
#plot trend for cross-sectional model and longitudinal model in same graph

# Define a range of ages to predict memory scores for different age points.
age_range <- data.frame(
  age_baseline = seq(
    from = min(merged_longitudinal_data_no_na_clean$age_baseline, na.rm = TRUE),
    to = max(merged_longitudinal_data_no_na_clean$age_baseline, na.rm = TRUE),
    length.out = 100))

# Add variables with default values
age_range <- age_range %>%
  mutate(
    phc_education = mean(merged_longitudinal_data_no_na_clean$phc_education, na.rm = TRUE),
    phc_diagnosis = factor(1, levels = levels(merged_longitudinal_data_no_na_clean$phc_diagnosis)),
    phc_race = factor(levels(merged_longitudinal_data_no_na_clean$phc_race)[1], 
                      levels = levels(merged_longitudinal_data_no_na_clean$phc_race)),
    phc_ethnicity = factor(levels(merged_longitudinal_data_no_na_clean$phc_ethnicity)[1], 
                           levels = levels(merged_longitudinal_data_no_na_clean$phc_ethnicity)),
    phc_sex = factor(levels(merged_longitudinal_data_no_na_clean$phc_sex)[1], 
                     levels = levels(merged_longitudinal_data_no_na_clean$phc_sex))
  )

# Make predictions using the cross-sectional model
age_range$predicted_phc_mem_cross_sectional <- predict(cross_sectional_model, newdata = age_range)

# Prepare data for longitudinal predictions
age_range_longitudinal <- age_range %>%
  mutate(time_since_baseline = 0)  # Assuming predictions at baseline

# Make predictions using the longitudinal model
age_range_longitudinal$predicted_phc_mem_longitudinal <- predict(model_lme, newdata = age_range_longitudinal, level = 0)

# Combine cross-sectional and longitudinal predictions
combined_data <- age_range %>%
  select(age_baseline, predicted_phc_mem_cross_sectional) %>%
  rename(predicted_phc_mem = predicted_phc_mem_cross_sectional) %>%
  mutate(model_type = "Cross-Sectional") %>%
  bind_rows(
    age_range_longitudinal %>%
      select(age_baseline, predicted_phc_mem_longitudinal) %>%
      rename(predicted_phc_mem = predicted_phc_mem_longitudinal) %>%
      mutate(model_type = "Longitudinal"))

# Plot predicted memory scores from both models
ggplot(combined_data, aes(x = age_baseline, y = predicted_phc_mem, color = model_type, linetype = model_type)) +
  geom_line(size = 1.2) +
  scale_color_manual(
    name = "Model Type",
    values = c("Cross-Sectional" = "blue", "Longitudinal" = "red"),
    labels = c("Cross-Sectional Model", "Longitudinal Model")) +
  scale_linetype_manual(
    name = "Model Type",
    values = c("Cross-Sectional" = "dashed", "Longitudinal" = "solid"),
    labels = c("Cross-Sectional Model", "Longitudinal Model")) +
  labs(
    title = "Predicted Memory Score Over Age by Cross-Sectional and Longitudinal Models",
    x = "Age (Years)",
    y = "Predicted Memory Score") +
  theme_minimal() 


```


```{r}
#plot for ad_drug Treatment effect

# Filter data for medication_category 1 (AD Drug)
ad_drug_data <- merged_longitudinal_data_no_na %>%
  filter(medication_category == 1)

# Filter data for other medications (not AD Drug)
other_drug_data <- merged_longitudinal_data_no_na %>%
  filter(medication_category != 1)


ggplot() + 
geom_line(data = other_drug_data, aes(x = time_since_baseline, y = phc_mem, color = "Other Medication", linetype = "Other Medication"), size = 1.2) +
geom_line(data = ad_drug_data, aes(x = time_since_baseline, y = phc_mem, color = "AD Drug", linetype = "AD Drug"), size = 1.2) +
scale_color_manual(
    name = "Medication Category",
    values = c("AD Drug" = "red", "Other Medication" = "blue")
  ) +
scale_linetype_manual(
    name = "Medication Category",
    values = c("AD Drug" = "solid", "Other Medication" = "dashed")
  ) +
  labs(
    title = "Treatment Effect Over Time by Medication Category",
    x = "Time Since Baseline (Years)",
    y = "Memory score",
    color = "Medication Category"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")


```

```{r}
# compare before ad drug medication treatment and after ad drug medication

# Arrange data by patient ID and time
treatment_data <- merged_longitudinal_data_no_na_clean%>%
  arrange(rid, time_since_baseline)

# Identify before and after treatment phases for each patient
treatment_data <- treatment_data %>%
  group_by(rid) %>%
  mutate(
    treatment_phase = case_when(
      # Before treatment: When medication_category is not yet 1
      medication_category != "1" & lag(medication_category, default = NA_character_) != "1" ~ "Before Treatment",
      # After treatment: When medication_category is 1
      medication_category == "1" ~ "After Treatment",
      # Default to NA if conditions aren't met
      TRUE ~ NA_character_
    )
  ) %>%
  ungroup()

# Filter data for before treatment and after treatment
before_treatment_data <- treatment_data %>% filter(treatment_phase == "Before Treatment")
after_treatment_data <- treatment_data %>% filter(treatment_phase == "After Treatment")

# Fit model for before treatment data
model_before <- lm(phc_mem ~ time_since_baseline, data = before_treatment_data)

# Fit model for after treatment data
model_after <- lm(phc_mem ~ time_since_baseline, data = after_treatment_data)

# Add predicted values to the original data for plotting
before_treatment_data <- before_treatment_data %>%
  mutate(predicted_phc_mem = predict(model_before, newdata = before_treatment_data))

after_treatment_data <- after_treatment_data %>%
  mutate(predicted_phc_mem = predict(model_after, newdata = after_treatment_data))


ggplot() +geom_point(data = before_treatment_data, aes(x = time_since_baseline, y = phc_mem, color = "Before Treatment"), size = 1.5, alpha = 0.6) +
  geom_point(data = after_treatment_data, aes(x = time_since_baseline, y = phc_mem, color = "After Treatment"), size = 1.5, alpha = 0.6) +
  geom_line(data = before_treatment_data, aes(x = time_since_baseline, y = predicted_phc_mem, color = "Before Treatment"), size = 1.2, linetype = "dashed") +
  geom_line(data = after_treatment_data, aes(x = time_since_baseline, y = predicted_phc_mem, color = "After Treatment"), size = 1.2, linetype = "solid") +
  scale_color_manual(
    name = "Treatment Phase",
    values = c("Before Treatment" = "blue", "After Treatment" = "red")
  ) +
  labs(
    title = "Comparison of Memory Score Trends Before and After Treatment",
    x = "Time Since Baseline (Years)",
    y = "Memory Score",
    color = "Treatment Phase"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")


```


# Model = baseline * cross-section + （visit age - baseline) * linear longitudinal
coef_longitudinal <- fixef(model_lme)  # For mixed effects model, use fixef() to get fixed effects
coef_cross_sectional <- coef(cross_sectional_model)

age_data <-merged_longitudinal_data_no_na_clean%>%
  select(time_since_baseline,phc_age_cognition)%>%
  mutate(age_baseline = round(phc_age_cognition - time_since_baseline,1))%>%
  mutate(predicted_phc_mem = coef_longitudinal['(Intercept)'] +
                             coef_cross_sectional['(Intercept)'] +
                             coef_cross_sectional['age_baseline'] * age_baseline +  # Cross-sectional effect of baseline age
                             coef_longitudinal['time_since_baseline'] * time_since_baseline)  # Longitudinal effect over time

# Fit a LOESS model to smooth the predicted values
loess_fit <- loess(predicted_phc_mem ~ time_since_baseline, data = age_data, span = 0.5)

# Add the smoothed values back to the data
age_data$smoothed_phc_mem <- predict(loess_fit)

# Plot the smoothed predicted values
ggplot(age_data, aes(x = time_since_baseline)) +
  # Plot observed data points
  geom_point(data = merged_longitudinal_data_no_na_clean, aes(y = phc_mem), color = "red", alpha = 0.6, size = 2) +
  # Plot smoothed predicted values as a line
  geom_line(aes(y = smoothed_phc_mem), color = "blue", size = 1.2) +
  labs(title = "Observed vs. LOESS Smoothed Predicted PHC Memory Scores",
       x = "Time Since Baseline (Years)",
       y = "PHC Memory Score") +
  theme_minimal() +
  theme(legend.position = "bottom")



