---
title: "adni_medication"
output: html_document
date: "2024-09-03"
---
```{r}
library(dplyr)
library(tidyr)
library(stringr)
library(tools)
library(purrr)
library(lubridate)
library(ADNIMERGE)
```



```{r}
medication = read.csv('/Users/viv/Desktop/data science/meds.csv')
adnimerge = read.csv('/Users/viv/Desktop/data science/adni_merge.csv')
# medication for backmeds
med34 = read.csv('/Users/viv/Desktop/data science/meds34.csv')
```

```{r}
# clean table adnimerge
cl_adnimerge <- adnimerge%>%
  janitor::clean_names()

#clean backmeds
backmed <- med34%>%
  janitor::clean_names()
#clean cmmed
medication<-medication%>%
  janitor::clean_names()%>%
  #drop empty 22411 rows
  filter(cmmed != -4)%>%
  mutate(cmmed = ifelse(cmmed == "", NA, cmmed))%>%
  drop_na(cmmed)%>%
  mutate(cmmed = str_to_title(cmmed))%>%
   # Remove double quotes from the strings
  mutate(cmmed = str_replace_all(cmmed, '"', "")) %>%
  # Remove single quotes from the strings
  mutate(cmmed = str_replace_all(cmmed, "'", "")) %>%

  mutate(cmmed = str_replace_all(cmmed, "\\[\\|194\\|\\]\\[\\|174\\|\\]", "")) 

#drop na column
medication <- medication[, colSums(is.na(medication)) < nrow(medication)]

  
```  
From dictionary, we identify that viscode2 is Translated visit code;  cmcont check if the medication continuing? 0=No; 1=Yes

```{r}
#clean letter a
cl_med_a<-medication%>%
  mutate(cmmed = case_when(
    str_detect(cmmed, "5% Saline Eye Drops") ~ "5% Saline Eye Drops",
    str_detect(cmmed, "Abs Sr. Plus") ~ "Abc Plus Senior",
    str_detect(cmmed, "Accopril") ~ "Accupril",
    str_detect(cmmed, "Acetaminophin|Acetamin") ~ "Acetaminophen",
    str_detect(cmmed, "Acetomenophen") ~ "Acetominophen",
    str_detect(cmmed, "Acetyl - L - Carnitine|Acetyl L Carnitine|Acetyl L- Canitine|Acetyl- L-Carnitine|Acetyl-L Carnitine") ~ "Acetyl L-Carnitine",
    str_detect(cmmed, "Acidolpheus|Acidophilus Pills|Acidophis") ~ "Acidophilus",
    str_detect(cmmed, "Acotonel|Actenol") ~ "Actonel",
    str_detect(cmmed, "Actose") ~ "Actos",
    #Acrobid might macrobid
    str_detect(cmmed,"Adalat Cc|Adalat Ex") ~"Adalat",
    str_detect(cmmed,"Advair Discus|Advair Disk") ~"Advair Diskus",
    #Advart might avodart
    str_detect(cmmed,"Aerobid Inhaler") ~"Aerobid",
    str_detect(cmmed,"Advil Es")~ "Advil",
    str_detect(cmmed,"Aggrenox Capsules Sa") ~ "Aggrenox",
    str_detect(cmmed,"Albuterol \\(Nebulizer\\)|Albuterol Nebulizer") ~ "Albuterol Inhaler",
    str_detect(cmmed,"Albuterol Sol" )~ "Albuterol Sulfate",
    str_detect(cmmed,"Aldara 5%")~ "Aldara",
    str_detect(cmmed,"Alenbronate|Alentdronate")~ "Alendronate",
    str_detect(cmmed,"Aledronate Sodium|Alendronate Sodium Usp")~ "Alendronate Sodium",
    str_detect(cmmed, "Align Cap") ~ "Align",
    str_detect(cmmed, "Alka Seltzer Plus|Alka Selzer Plus|Alka-Seltzer Plus Cold") ~ "Alka-Seltzer Plus",
    str_detect(cmmed,"Allegra \\(Fexofenadine\\)")~"Allegra",
    str_detect(cmmed, "Allergen Injections|Allergy Injection") ~ "Allergy Shots",
    str_detect(cmmed,"Allevert")~"Alleve",
    str_detect(cmmed,"Allopurinal")~"Allopurinol",
    str_detect(cmmed,"Alpha-Lipoic Acid")~"Alpha Lipoic Acid",
    str_detect(cmmed,"Alphagan P \\(Brimonidine Tartrate Ophthalmic Solution\\)|Alphagan P.o 15%") ~"Alphagan P",
    str_detect(cmmed, "Alphagan 0.15%") ~ "Alphagan",
    str_detect(cmmed,"Alprazalam|Alprazolm") ~"Alprazolam",
    str_detect(cmmed,"Amiordarone")~"Amiodarone",
    str_detect(cmmed,"Amitryptiline|Amittriptyline")~"Amitriptyline",
    str_detect(cmmed,"Amlod/Benaze") ~"Amlodipine/Benzapine",
    str_detect(cmmed,"Amoxicilin|Amoxicillan|Amoxicliin|Amoxil")~"Amoxicillin",
    str_detect(cmmed,"Androgel 1 %")~"Androgel 1%",
    str_detect(cmmed, "Antacid Tablet")~"Antacid Tablets",
    str_detect(cmmed,"Antibiotic \\(Name Unknown\\)")~"Antibiotic",
    str_detect(cmmed,"Antihistamine")~"Antihistamines",
    str_detect(cmmed,"Arenesp")~"Aranesp",
    str_detect(cmmed,"Aricept For Memory Impairment|Aricpet")~"Aricept",
    str_detect(cmmed,"Arimedex")~"Arimidex",
    str_detect(cmmed,"Armour|Armours Thyroid")~"Armour Thyroid",
    str_detect(cmmed,"Asa \\(Baby Asprin\\)|Asa \\(Baby\\)|Asa-Baby Asprin")~"Asa",
    str_detect(cmmed,"Aspirine|Aspril|Asprin|Aspririn")~"Aspirin",
    str_detect(cmmed,"Astelin Nasal Spray|Astelin Nose Spray")~"Astelin",
    str_detect(cmmed,"Atacand/Hctz")~"Atacand Hct",
    str_detect(cmmed,"Atenolol La|Atenol")~"Atenolol",
    str_detect(cmmed,"Atenolol/ Chlorthal|Atenolol+Chlorthalidine")~"Atenolol/Chlorthalidine",
    str_detect(cmmed,"Atrovastatin") ~"Atorvastatin",
    str_detect(cmmed,"Atrovant .06") ~"Atrovent 0.06%",
    str_detect(cmmed, "Azithronycin|Azitromycin") ~ "Azithromycin",

    
    str_detect(cmmed, "vit c") ~ "Vitamin C",
   
    TRUE ~ cmmed
  ))

```

```{r}
# Corrected code with str_detect used properly
cl_med_b <- cl_med_a %>%
  mutate(cmmed = case_when(
    str_detect(cmmed, "B 100") ~ "B-100",
    str_detect(cmmed, "B 12") ~ "B-12",
    str_detect(cmmed, "B Complex|B- Complex|Complex B") ~ "B-Complex",
    str_detect(cmmed, "B Complex Vitamin|B Complex/Vitamin|B Complex Vitamins") ~ "B-Complex Vitamin",
    str_detect(cmmed, "B-Complex 50") ~ "B-50 Complex",
    str_detect(cmmed, "Benicar Hctz") ~ "Benicar Hct",
    str_detect(cmmed, "Beta Carotine|Beta-Carotene") ~ "Beta Carotene",
    str_detect(cmmed, "Bethamethasone") ~ "Betamethasone",
    str_detect(cmmed, "Bethamethasone Valerate") ~ "Betamethasone Valerate",
    str_detect(cmmed, "Bisopril \\(Hctz\\)|Bisopropolol/Hctz") ~ "Bisoprolol Hctz",
    str_detect(cmmed, "Black Cohash") ~ "Black Cohosh",
    str_detect(cmmed, "Bone Up") ~ "Boneup",
    str_detect(cmmed, "Brimonidine Tartate") ~ "Brimonidine Tartrate",
    str_detect(cmmed, "Buproprion") ~ "Bupropion",
    str_detect(cmmed, "Buproprion Er") ~ "Bupropion Er",
    str_detect(cmmed, "Buproprion Hcl") ~ "Bupropion Hcl",
    str_detect(cmmed, "Buproprion SR") ~ "Bupropion SR",
    TRUE ~ cmmed
  ))

```

```{r}
cl_med_c <- cl_med_b %>%
  mutate(cmmed = case_when(
    str_detect(cmmed,"Ca, Mg, K")~"Ca + Mg + K",
    str_detect(cmmed, regex(
      "Ca W/ Vitamin D|Calcium & D|Calcium & Vitamin D|Calcium \\+ D|Calcium \\+ Vit D|Calcium \\+ Vit\\. D|Calcium \\+D|Calcium \\+Vit D|Calcium / Vitamin D|Calcium And Vitamin D|Calcium Vit D|Calcium/Vit D|Calcium/Vitamin D|Calcuim \\+ Vitamin D|Calcuim And Vitamin D|Calcuim Plus Vitamin D|Calcuim With Vitamin D|Calcium-Vitamin D|Ca/D|Calcium /Vitamin D|Calcium D|Calcium Plus Vit D|Calcium Plus Vitamin D|Calcium Plus D|Calcium W/ Vit D|Calcium W/ Vitamin D|Calcium W/D|Calcium W/Vit D|Calcium W/Vitamin D|Calcium With D|Calcium With Vit D|Calcium With Vit. D|Calcium With Vitamin D|Calcium Wtih Vit D|Calcium\\+D|Calcium\\+Vitamin D|Calicum \\& Vitamin D|Calcium/ Vitamin D|Calcium/D", 
      ignore_case = TRUE)) ~ "Calcium + Vitamin D",
    str_detect(cmmed, regex("Calcium \\(500mg\\) \\+ Vitamin D \\(125iu\\)", ignore_case = TRUE)) ~ "Calcium (500 Mg) + Vitamin D3 (125 Iu)",
    str_detect(cmmed,"Calcium And Magnesium|Cal Mag|Cal Mag T|Cal-Mag|Calcium Mag|Calcium Magnesium|Calcium With Magnesium|Calcium-Magnesium|Calcium\\&Magnesium Suppl.|Calcium/Magnesium|Calcium/Mg")~"Calcium + Magnesium",
    str_detect(cmmed,"Calcium With Vitamind D And Magnesium")~"Calcium, Magnesium And Vitamin D",
    str_detect(cmmed,"Calcium/Zing/Magnesium Supplement")~"Calcium, Magnesium, Zinc",
    str_detect(cmmed,"Calcium Carbonate Vitamin D|Calcium Carbonate W/ Vitamin D|Calcium Carbonate W/Vitamin D|Calcium Carbonate/Vitamin D")~"Calcium Carbonate + Vitamin D",
    str_detect(cmmed,"Calcium, Magnesium \\& Zinc|Calcium, Zinc, Magnesium")~"Calcium, Magnesium, Zinc",
    str_detect(cmmed,"Calcium Cit")~"Calcium Citrate",
    str_detect(cmmed,"Calcium Citrate \\+D|Calcium Citrate Plus D|Calcium Citrate With Vit D|Calcium Citrate With Vitamin D|Calcium Citrate Wvitamin D Nos|Calcium Citrate/Vit D|Calcium Citrate/Vitamin D")~"Calcium Citrate +Vitamin D",
    str_detect(cmmed,"Calcium Citrate Magnesium|Calcium/Magnesium Citrate")~"Calcium Citrate With Magnesium",
    str_detect(cmmed,"Calicum")~"Calcium",
    str_detect(cmmed, regex("Calcium 600 - D|Calcium 600 D|Calcium 600 \\+ D|Calcium 600\\+D", ignore_case = TRUE)) ~ "Calcium 600 Plus Vitamin D",
      str_detect(cmmed,"Caltrate With Vitamin D|Caltrate-D|Caltrate Plus")~"Caltrate D",
    str_detect(cmmed,"Carbidopalevodopa")~"Carbidopa/Levodopa",
    str_detect(cmmed,"Cartia Xl")~"Cartia Xt",
    str_detect(cmmed, "Celexia")~"Celexa",
    str_detect(cmmed,"Centralvite")~"Central Vite",
    str_detect(cmmed,"Centrum A To Zinc")~"Centrum A-Z",
    str_detect(cmmed,"Centravite Sr|Centrum Snior")~"Centrum Senior",
    str_detect(cmmed,"Centrum-Silver|Centrum Sliver|Centrum Silver 2000")~"Centrum Silver",
    str_detect(cmmed,"Centrum Silver Mvi|Centrum Silver Multivitamin|Centrum Silver, Multivitamins|Centrum Silver- Mvi")~"Centrum Silver Multi-Vitamin",
    str_detect(cmmed,"Cephalfrin")~"Cephalexin",
    str_detect(cmmed,"Cetiricine Hcl|Cetirizine Hcl")~"Cetirizine Hydrochloride",
    str_detect(cmmed,"Choline Bititrate")~"Choline Bitartrate",
    str_detect(cmmed,"Chondroiton/Glucosamine|Chondroitin With Glucosamine|Chondroitin Wglucosamine|Chondroitin Glucosamine")~"Chondroitin/Glucosamine",
    str_detect(cmmed,"Chromium Picolonate")~"Chromium Picolinate",
    str_detect(cmmed,"Ciproflaxacin Hcl")~"Ciprofloxacin Hcl",
    str_detect(cmmed,"Citalopran")~"Citalopram",
    str_detect(cmmed,"Citracal \\+ Vit D|Citracal \\+D|Citracal D\\+|Citrical With D|Citrical\\+Vitamin D")~"Citracal + D",
    str_detect(cmmed,"Citrocel")~"Citrucel",
    str_detect(cmmed,"Claratin|Claritan")~"Claritin",
    str_detect(cmmed,"Climera")~"Climara",
    str_detect(cmmed,"Clindamyacim")~"Clindamycin",
    str_detect(cmmed,"Clobetasol Proprionate|Clobetasone Propionate")~"Clobetasol Propionate",
    str_detect(cmmed,"Clonazepan|Clonezepam")~"Clonazepam",
    
    str_detect(cmmed, regex("Combigan 0\\.2/0\\.5%", ignore_case = TRUE)) ~ "Combigan 0.2% / 0.5% Oph",
    str_detect(cmmed, regex("Congugated Estrogen", ignore_case = TRUE)) ~ "Conjugated Estrogen",
    str_detect(cmmed, regex("Coq-10|Coq 10|Coq\\=10|Coq10 Enzyme|Coq10|Coenzyme Q10 \\(Coq10\\)|Coenzime Q10|Coenzyme Q|Coenzyme Q 10|Coenzyme Q-10|Co-Enzyme Q|Co-Enzyme Q-10|Co-Enzyme Q10|Co-Q|Co-Q-10|Co-Q10|Co Q.10|Co Q10|Co Q10 Enzyme", ignore_case = TRUE)) ~ "Coenzyme Q10",
    str_detect(cmmed,"Cortisone Injections")~"Cortisone Injection",
    str_detect(cmmed, regex("Cozar", ignore_case = TRUE)) ~ "Cozaar",
    str_detect(cmmed, regex("Cran Actin", ignore_case = TRUE)) ~ "CranActin",
    str_detect(cmmed, regex("Cranberry \\+C", ignore_case = TRUE)) ~ "Cranberry + C",
    TRUE ~ cmmed
  ))

```

```{r}
cl_med_def <- cl_med_c %>%
  mutate(cmmed = case_when(
    str_detect(cmmed,"Darvocet N-100")~"Darvocet-N 100",
    str_detect(cmmed,"Decaffeinated Greentea")~"Decaffeinated Green Tea",
    str_detect(cmmed,"Depahote")~"Depakote",
    str_detect(cmmed,"Diazepan")~"Diazepam",
    str_detect(cmmed,"Diclofen Sodium|Diclofenac Sod")~"Diclofenac Sodium",
    str_detect(cmmed,"Diltiazam")~"Diltiazem",
    str_detect(cmmed,"Diltiazem Hcl")~"Diltiazem Hydrochloride",
    str_detect(cmmed,"Diovan/ Hct")~"Diovan Hct",
    str_detect(cmmed,"Diphenoxylateatropine")~"Diphenoxylate/Atropine",
    str_detect(cmmed,"Dipyridimole")~"Dipyridamole",
    str_detect(cmmed,"Doxycycle")~"Doxycycline",
    str_detect(cmmed,"Duo-Neb")~"Duoneb",
    str_detect(cmmed,"Dynacirc, Cr")~"Dynacirc Cr",
    str_detect(cmmed,"Econopred Plus")~"Econopred+",
    str_detect(cmmed,"Ec Asa|Ec Aspirin|Ec-Asa|Ecasa")~"Ec-Aspirin",
    str_detect(cmmed,"Effexor Xl|Effexor \\(Xr\\)|Effexor - Xr|Effexor Er")~"Effexor Xr",
    str_detect(cmmed,"Elidel 1 % Cream")~"Elidel 1% Cream",
    str_detect(cmmed,"Elocom 0.1%|Elocom Cream")~"Elocom 0.1 % Cream",
    str_detect(cmmed,"Enablex Sr")~"Enablex Er",
    str_detect(cmmed,"Enteric Coated Acetlysalicylic Acid|Enteric Coated Acetysalicylic Acid")~"Enteric Coated Acetylsalicylic Acid",
    str_detect(cmmed,"Enteric-Coated Aspirin")~"Enteric Coated Aspirin",
    str_detect(cmmed,"Epa")~"Epa Fish Oil",
    str_detect(cmmed,"Eph-Dha|Epa-Dpa Liquid|Epa/Dha")~"Epa, Dpa Liquid-Flavored Fish Oil",
    str_detect(cmmed,"Erythromyan")~"Erythromycin",
    str_detect(cmmed,"Erythromycin Opthalmic Ointment")~"Erythromycin Ophthalmic Ointment",
    str_detect(cmmed,"Ester-C")~"Ester C",
    str_detect(cmmed,"Excedrine")~"Excedrin",
    str_detect(cmmed,"Exelon Parch|Exelon - Patch|Exelon Patch|Exelon 10cm2")~"Exelon",
    str_detect(cmmed,"Exforgen")~"Exforge",
    str_detect(cmmed,"Experimental Drug For Htn")~"Experimental Htn Drug",
    str_detect(cmmed,"Extra Stength Tylenol|Extra-Strength Tylenol")~"Extra Strength Tylenol",
    str_detect(cmmed,"Ezetimbe|Eztimibe")~"Ezetimibe",
    str_detect(cmmed,"Fem Ring")~"Femring",
    str_detect(cmmed,"Fentanyl Patch")~"Fentanyl Transdermal Patch",
    str_detect(cmmed,"Feosol Iron|Fesol")~"Feosol",
    str_detect(cmmed,"Fergon Iron")~"Fergon",
    str_detect(cmmed,"Feso4")~"Ferrous Sulfate",
    str_detect(cmmed,"Fexofenidine")~"Fexofenadine",
    str_detect(cmmed,"Fiber-Con")~"Fibercon",
    str_detect(cmmed,"Finesteride|Finesceride")~"Finasteride",
    str_detect(cmmed,"Fish Body Oils 1000")~"Fish Body Oil 1000",
    str_detect(cmmed,"Fish Oil (Omega-3)|Fish Oil / Omega 3|Fish Oil - Omega 3|Fish Oil Omega 3|Fish Oil/ Omega 3")~"Fish Oil + Omega 3",
    str_detect(cmmed,"Fishoil")~"Fish Oil",
    str_detect(cmmed,"Fish Oil Caps")~"Fish Oil Capsules",
    str_detect(cmmed,"Fish Oil - Omega 3 Fatty Acids|Fish Oil /Omega-3 Fatty Acids")~"Fish Oil + Omega 3 Fatty Acids",
    str_detect(cmmed,"Flackseed Oit|Flackseed Oil|Flak Seed Oil|Flax Oil|Flax Seed Oil")~"Flaxseed Oil",
    str_detect(cmmed,"Flax Seed")~"Flaxseed",
    str_detect(cmmed,"Flexerall|Flexaril|Flexoril")~"Flexeril",
    str_detect(cmmed,"Flonaise")~"Flonase",
    str_detect(cmmed,"Flunisolide Nasal Spray 0.025%|Flunisolide|Flunisolide 0.025%")~"Flunisolide Nasal Spray",
    str_detect(cmmed,"Flut Shot")~"Flu Shot",
    str_detect(cmmed,"Fluticaseon Proprionate|Fluticasone Propinate")~"Fluticasone Propionate",
    str_detect(cmmed,"Follic Acid")~"Folic Acid",
    str_detect(cmmed,"Foltex")~"Foltx",
    str_detect(cmmed,"Formotoral Fumarate|Fomoterol Inhaler")~"Formoterol Fumarate",
    str_detect(cmmed,"Forward Multimvitamins")~"Forward Multivitamins",
    str_detect(cmmed,"Fosamax D|Fosamax \\+D|Fosamax/D|Fosomax \\+ D")~"Fosamax Plus D",
    str_detect(cmmed,"Fosamex|Fosomax|Foxamax")~"Fosamax",
    str_detect(cmmed,"Fosomax Weekly")~"Fosamax Weekly",
    TRUE ~ cmmed
  ))


```

```{r}
cl_med_ghi <- cl_med_def %>%
  mutate(cmmed = case_when(
str_detect(cmmed,"Gabbapentin")~"Gabapentin",
str_detect(cmmed,"Garlic Tablet|Garlic Tablets")~"Garlic Tabs",
str_detect(cmmed,"Gas Ex")~"Gas-X Ex",
str_detect(cmmed,"Gas X")~"Gas-X",
str_detect(cmmed,"Gelatin Caps|Gelatin")~"Gelatin Capsules",
str_detect(cmmed,"Gemfibrazol")~"Gemfibrozil",
str_detect(cmmed,"Gentamycin")~"Gentamicin",
str_detect(cmmed,"Gingkko Biloba")~"Gingko Biloba",
str_detect(cmmed,"Ginkgo Bilboba|Ginko Balboa|Ginko Biloba")~"Ginkgo Biloba",
str_detect(cmmed,"Ginko")~"Ginkgo",
str_detect(cmmed,"Ginko Biloba Plus Tablet")~"Ginkgo Biloba Plus Tablet",
str_detect(cmmed,"Glucosamide Chondrotin|Glucosamin And Chondroitin|Glucosamin Chondroitin|Glucosamine  Chondroitin|Glucosamine \\+ Chondroitin|Glucosamine / Chondroitin|Glucosamine /Chondroitin|Glucosamine And Chondriton|Glucosamine And Chondroitin|Glucosamine Chondrointine|Glucosamine Chondroitin|Glucosamine Chondroiton|Glucosamine Chondrotin|Glucosamine Chrondroiten|Glucosamine Chrondroitin|Glucosamine-Chondroitin|Glucosamine/ Chondroitin|Glucosamine/ Chondrotin|Glucosamine/Chondriotin|Glucosamine/Chondroiten|Glucosamine/Chondroitin|Glucosamine/Chondrotin|Glucosamine/Chrondrotin|Glucosamine/Condroitin|Glucosamine Plus Chondroitin")~"Glucosamine & Chondroitin",
str_detect(cmmed,"Glucosimine|Glucosomine")~"Glucosamine",
str_detect(cmmed,"Glucosamine And Msn|Glucosamine Msm|Glucosamin/Msm|Glucosamine With Mgs|Glucosamine/Msn|Glucosamine With Msm|Glucosamine W/Msn")~"Glucosamine/Msm",
str_detect(cmmed,"Glucosamine//+Chondroiton//+Msm")~"Glucosamine/Chondroitin/Msm",
str_detect(cmmed,"Glucosamine Sulphate")~"Glucosamine Sulfate",
str_detect(cmmed,"Glucosamine Hc1")~"Glucosamine Hcl",
str_detect(cmmed,"Glucotrol X L|Glucostrol Xl")~"Glucotrol Xl",
str_detect(cmmed,"Glybor/Metfrm|Glyburidemetformin")~"Glyburide/Metformin",
str_detect(cmmed,"Guafenesin|Guaifeneson|Guiafenesin|Guaifensin")~"Guaifenesin",
str_detect(cmmed,"Glucosamine Chondrioten Sulfate|Glucosamine And Chondroiton Sulfate")~"Glucosamine Chondroitin Sulfate",
str_detect(cmmed,"Hawthorne")~"Hawthorn",
str_detect(cmmed,"Hawthorne Berries")~"Hawthorn Berries",
str_detect(cmmed,"Hctz|Htcz|Hydorchlorothiazide|Hctz (Hydrochlorothiazide)|Hydrochlorothiazide (Hctz)|Hydrochlorathiazide|Hydrochlorithiazide|Hydrochlorot|Hydrochlorothyazide|Hydrochlorthiazide|Hydrocholorthiazide|Hydroclhlorazide|Hydroclorot")~"Hydrochlorothiazide",
str_detect(cmmed,"Herblax")~"Herb-Lax",
str_detect(cmmed,"Humera")~"Humira",
str_detect(cmmed,"Hyalgin")~"Hyalgan",
str_detect(cmmed,"Hyalurenic Acid")~"Hyaluronic Acid",
str_detect(cmmed,"Hydorcodone")~"Hydrocodone",
str_detect(cmmed,"Hydrocortisone Cream 1%")~"Hydrocortisone 1% Cream",
str_detect(cmmed,"Hydrocodone - Apap|Hydrocodone/Apap|Hydrocodone/Atat|Hydrocodone W/ Acetominophen|Hydrocodoneacetaminophen")~"Hydrocodone Acetaminophen",
str_detect(cmmed,"Hydromorphine")~"Hydromorphone",
str_detect(cmmed,"Hydroxizine|Hydroxydine|Hydrozyine")~"Hydroxyzine",
str_detect(cmmed,"Hyoscyamin")~"Hyoscyamine",
str_detect(cmmed,"Hyoscyamin Sltb")~"Hyoscyamine Sltb",
str_detect(cmmed,"Hyrodomorphone Hcl")~"Hydromorphone Hcl",
str_detect(cmmed,"I Cap|I Caps|I-Cap")~"I-Caps",
str_detect(cmmed,"I-Caps (Arends Formula)")~"I-Caps (Areds Formula)",
#ad drug
str_detect(cmmed,"Galantamine (Razadyne)|Galantamine Er|Galantamine Hydrobromide")~"Galantamine",
str_detect(cmmed,"Revistagmine")~"Rivastigmine",
str_detect(cmmed,"Rezadyne")~"Razadyne",
str_detect(cmmed,"Namenda Hcl|Nameda")~"Namenda",
str_detect(cmmed,"Memantine Hcl|Mematine|Mementine")~"Memantine",
#anti-depression drug
str_detect(cmmed,"Mirtazapine Tablet")~"Mirtazapine",
str_detect(cmmed,"Prozac Fluoxetine Hcl|Fluoxetine Hcl")~"Fluoxetine",
str_detect(cmmed,"Sertraline Hcl")~"Sertraline",
str_detect(cmmed,"Trazodone Hcl")~"Trazodone",
str_detect(cmmed,"Paroxetine Hydrochloride|Paroxetine Hcl")~"Paroxetine",
str_detect(cmmed,"Paxil Cr")~"Paxil",
str_detect(cmmed,"Venlafaxine Hydrochloride Xr")~"Venlafaxine",
str_detect(cmmed,"Budaprion (Wellbutrin)|Wellbutrin Sr|Wellbutrin Xl|Wellbutrin Xr|Wellbutrin-Xl|Wellbutrinxl")~"Wellbutrin",
str_detect(cmmed,"Amitriptyline Hcl")~"Amitriptyline",
str_detect(cmmed,"Bupropion Er|Bupropion Hcl|Bupropion Xl")~"Bupropion",
str_detect(cmmed,"Citalopram Hbz|Citalopram Hydrobromide")~"Citalopram",
str_detect(cmmed,"Effexor Xr")~"Effexor",
TRUE ~ cmmed
  ))

```
Then we identify ad-specific drugs and anti-depression drugs,  after that, we add a new column called medication category, where ad-drugs = 1, antidep-drugs = 2, other drugs = 0, so that we can keep original cmmed while knowing the category of each medication from each visit
```{r}
# AD-specific drugs
ad_drugs <- c(
  "Aricept", "Donepezil", "Donepezil Hcl", "Exelon", "Razadyne",
  "Rivastigmine", "Galantamine", "Namenda", "Memantine",
  "Hydergine Ergoloid Mesylates", "Ergoloid Mesylate"
)

# Anti-depression drugs
antidep_drugs <- c(
  "Fluoxetine", "Prozac", "Sertraline", "Lexapro", "Escitalopram Oxalate",
  "Paxil", "Paroxetine", "Citalopram", "Celexa", "Cymbalta",
  "Venlafaxine", "Effexor", "Bupropion", "Wellbutrinxl",
  "Mirtazapine", "Remeron", "Trazodone", "Desyrel",
  "Amitriptyline", "Nortriptyline", "Imipramine", "Tofranil Pm","Zoloft","Sertraline"
)


cl_med <- cl_med_ghi %>%
  group_by(ptid, viscode, phase) %>%
  # Create a concatenated list of all unique medications per visit
  mutate(all_medications = paste(unique(cmmed), collapse = ", ")) %>%
  # Determine medication category for each visit
  mutate(
    medication_category = case_when(
      any(cmmed %in% ad_drugs) & any(cmmed %in% antidep_drugs) ~ "1|2",  # If both AD and antidepressant drugs are present
      any(cmmed %in% ad_drugs) ~ "1",                                   # 1 if any of the AD drugs are present
      any(cmmed %in% antidep_drugs) ~ "2",                              # 2 if any of the antidepressant drugs are present
      TRUE ~ "0"                                                        # 0 if neither is present
    )
  ) %>%
  ungroup() %>%
  select(rid, phase, viscode, visdate, cmmed, medication_category, all_medications, everything())

# Count the number of records for each medication category
cl_med %>%
  group_by(medication_category) %>%
  count()

```

```{r}

medication_keys <- cl_med %>% 
  select(rid, phase, viscode) %>% 
  distinct()
backmed_keys <- backmed %>% 
  select(rid, phase, viscode) %>% 
  distinct()

# Find the intersection of keys
common_keys <- intersect(medication_keys, backmed_keys)
  
# add keymeds from backmeds with medication table 
medication_with_keymed <- cl_med %>%
  full_join(backmed  %>% 
              select(rid, phase, viscode, keymed), 
            by = c("rid" = "rid", "phase" = "phase", "viscode" = "viscode"))%>%
            select(rid, phase, viscode, keymed,everything())

# add medication category based on keymed
cl_med_na <- medication_with_keymed %>%
  filter(is.na(medication_category))%>%
  mutate(
    keymed_codes = str_split(keymed, "\\|"),
    # Assign medication_category based on keymed_codes
    medication_category = case_when(
      map_lgl(keymed_codes, ~ any(.x %in% c("1", "2", "3", "4", "5"))) ~ 1,
      map_lgl(keymed_codes, ~ any(.x == "6")) ~ 2,
      map_lgl(keymed_codes, ~ any(.x == "7")) ~ 3, # assign other behavior medication to new category 3
      TRUE ~ 0
    )) %>%
  select(-keymed_codes)

# Update medication_with_keymed to include new categories and resolve errors in type mismatch
medication_with_keymed_updated <- medication_with_keymed %>%
  left_join(
    cl_med_na %>%
      select(rid, phase, viscode, cmmed, medication_category),
    by = c("rid", "phase", "viscode", "cmmed"),
    suffix = c("", "_new")
  ) %>%
  mutate(
    # Convert both columns to character before using coalesce
    medication_category = coalesce(as.character(medication_category_new), as.character(medication_category))
  ) %>%
  select(-medication_category_new)

# View the updated table
medication_with_keymed_updated%>%
    group_by(medication_category) %>%
  count()


```

One issue here is that when keymed like 1|6, or 6|7, medication category only show one of them, but without further information on keymed, i can only fill out NA this way.
```{r}
# Add keymed with NA based on cmmed
#1=Aricept;2=Cognex;3=Exelon;4=Namenda;5=Razadyne;
#1-5: ad-drugs 6=Anti-depressant medication;7=Other behavioral medication;0=None of the above
med_code_mapping <- c(
  "Aricept" = 1,
  "Cognex" = 2,
  "Exelon" = 3,
  "Namenda" = 4,
  "Razadyne" = 5)

# Function to map medications to codes
map_meds_to_codes <- function(cmmed_list, med_category_list) {
  codes_present <- c()

  cmmed_list_lower <- tolower(cmmed_list)
  
  # Map medications with medication_category == 1
  for (med in names(med_code_mapping)) {
    med_lower <- tolower(med)
    
    indices <- which(str_detect(cmmed_list_lower, fixed(med_lower)))
   
    if (length(indices) > 0 && any(med_category_list[indices] == 1)) {
      codes_present <- c(codes_present, med_code_mapping[med])
    }
  }
  
  # Add '6' if any medication_category == 2
  if (any(med_category_list == 2)) {
    codes_present <- c(codes_present, 6)
  }
  
  # If no codes are present, add '0'
  if (length(codes_present) == 0) {
    codes_present <- c(0)
  }
  
  # Return the codes as a string separated by '|', sorted and unique
  return(paste(sort(unique(codes_present)), collapse = "|"))
}

visit_medication <- medication_with_keymed_updated %>%
  group_by(rid, phase, viscode) %>%
  summarise(
    cmmed_list = list(cmmed),
    med_category_list = list(medication_category),
    .groups = 'drop'
  ) %>%
  mutate(
    keymed = pmap_chr(
      list(cmmed_list, med_category_list),
      map_meds_to_codes
    )
  ) %>%
  ungroup()%>%
  select(-cmmed_list, -med_category_list)

medication_with_keymed_updated <- medication_with_keymed_updated %>%
  left_join(visit_medication, by = c("rid", "phase", "viscode"))


medication_with_keymed_updated%>%
  group_by(keymed.y)%>%
  count()


```

I've updated keymed based on the dictionary of keymed, one potential issue here is ad-drugs that doesn't exist in the keymed dictionary will assign 0 to keymed if there are no other medications contributing to keymed.

```{r}
#check exploratory data for medication with keymed
medication_with_keymed%>%
  group_by(keymed)%>%
  count()

#Exploratory Data
medication%>% 
  group_by(cmreason)%>%
  count()
medication%>%
  group_by(phase)%>%
  count()
backmed%>%
  group_by(phase)%>%
  count()

medication%>%
  filter(rid %in% c(6463,6535))
```


```{r}
#metformin 
metformin <-cl_med_ghi%>%
  filter(str_detect(cmmed, regex("metformin|Glucophage", ignore_case = TRUE)))

View(metformin)

#check metformin rid with the article
metformin_rid = c(339, 1408, 1423, 2063, 2072, 2193, 2263, 4077, 4432, 4653, 4675, 4689, 4765, 4803, 4919, 6463, 6535)

metformin%>%
  filter(rid %in% metformin_rid)

# Install ADNIMERGE from the local source file
#reccmeds<-ADNIMERGE::reccmeds

# check metformin from r package
#reccmeds%>%
#  filter(RID %in% metformin_rid)

na_keymed_rows <- visit_medication %>%
  filter(is.na(keymed))
```



```{r}
matched_entries <- metformin %>%
  inner_join(cl_adnimerge, by = "ptid")%>%
  select(viscode.y, examdate, everything())

matched_entries <- matched_entries%>%
  group_by(ptid,viscode.y,examdate)


View(matched_entries)
```


```{r}
# Load the cognitive marker data
cognitive_marker <- read.csv("/Users/viv/adni_medication/ADSP_ADNI_Cognition_Dec2023/ADSP_PHC_COGN_Dec2023.csv")

# Clean column names using janitor::clean_names() for consistency
cl_cognitive_marker <- cognitive_marker %>%
  janitor::clean_names()

# Add keymed and medication category by joining with the medication dataset

# add keymed and medication category
cl_cognitive_marker <- cl_cognitive_marker %>%
  left_join(medication_with_keymed_updated %>% 
              select(rid, viscode2, phase, keymed.y,all_medications,medication_category,visdate),
            by = c("rid", "viscode2", "phase"))%>%
  select(rid, viscode2, phase, examdate, visdate,keymed.y, all_medications,medication_category,everything())


cognitive_marker_distinct <- cl_cognitive_marker %>%
  distinct(rid, viscode, phase, .keep_all = TRUE)
cl_cognitive_marker%>%
  group_by(medication_category)%>%
  count()
```



```{r}

# Convert examdate and visdate columns to Date class
cognitive_marker_date <- cognitive_marker_distinct %>%
  mutate(examdate = ymd(examdate),visdate = ymd(visdate))


```

